---
title: "Web Scraping Assignment 2"
author: "Nawal Zehra Hasan"
date: "12/05/2021"
output: html_document
---


Prompt: Find an api that returns with a JSON. You have to prove that you know how to deal with a list. 
Write a function(s) that process one batch of data and then move to the next data. (lapply, rbindlist). Show your data in table format(head is enough) and make a few(2-3) ggplot which will give some insight of your dataset.   


```{r, echo=TRUE, message=FALSE}
#cleaning the environment
rm(list = ls())
```


```{r, echo=TRUE, message=FALSE}
#loading relevant libraries
library(jsonlite)
library(httr)
library(rvest)
library(data.table)
library(ggplot2)
library(tidyverse)
library(forcats)
```
ABOUT THE API

To complete this task, I chose the offical US Government's source for fuel economy API information to scrape data. The API information can be found [*here*](https://www.fueleconomy.gov/feg/ws/). This was my choice as it gave a variety of options to choose from about vehicles such as vehicle make, fuel prices, emissions, model years and much more. I chose emissions for a few vehicle IDs(provided in the csv on the website) which returned a data frame with the variables listed below. This API was publicly available and anyone can access it without any sort of registration or private key/token. 

Below is the meaning for each variable for Emissions dataframe that I will create using the API
* EmissionsInfo
* efid: engine family ID
* id: vehicle record ID (links emission data to the vehicle record)
* salesArea: EPA sales area code
* score: EPA 1-10 smog rating for fuelType1
* scoreAlt: EPA 1-10 smog rating for fuelType2
* smartwayScore: SmartWay Code
* standard: Vehicle Emission Standard Code
* stdText: Vehicle Emission Standard

```{r, echo=TRUE, results='hide'}
#emission records for specific vehicle
#using base url
base_url <- 'https://www.fueleconomy.gov/ws/rest/'
#creating info url
info_url <- 'vehicle/emissions/'
#adding ID
id <- '40593'
full_url <- base::paste0(base_url, info_url, id)
#call the API
api_call <- GET(full_url)  
api_call$status_code
Emissions_list <- fromJSON(rawToChar(api_call$content), flatten = TRUE)
df <- Emissions_list$emissionsInfo
df
```


```{r, echo=TRUE, results='hide'}
#emission records for 4 vehicles
base_url <- 'https://www.fueleconomy.gov/ws/rest/'
info_url <- 'vehicle/emissions/'
ids <- c('41118','41090', '40638','19935')
full_url <- paste0(base_url, info_url, ids)
```



```{r, echo=TRUE, results='hide'}
#function to get multiple vehicles emissions
Call_api <- function(api){
  api_call <- GET(api)
  api_call$status_code
  Emissions_list <- fromJSON(rawToChar(api_call$content), flatten = TRUE)
  df <- Emissions_list$emissionsInfo
}
```


```{r}
#using lapply and rbind 
df_list <-lapply(full_url,Call_api)
final_df <- rbindlist(df_list) 
final_df  
head(final_df)
```


```{r, fig.align="center"}
#Visualizations

#no. of vehicles with given IDs
v1 <- ggplot(final_df, aes(x= fct_infreq(id), fill = id)) +
  geom_bar() +
  xlab('Vehicle Id') +
  ylab('No. of vehicles') + 
  ggtitle('Number of Vehicles with Given IDs') +
  scale_fill_manual(values = c("19935" = "steelblue", 
                               "41090" =  "skyblue", 
                               "41118" = "lightblue", 
                               "40638" = "slategray2"))+
   theme(plot.title = element_text(face = "bold", hjust = 0.5))
v1
```


```{r, fig.align="center"}
#Emissions Score by Sales Area
v2 <- ggplot(final_df) +
  aes(x = score, fill = salesArea, colour = salesArea) +
  geom_bar() +
  scale_fill_manual(
    values = c(`3` = "#29749E",
               `7` = "#122243")) +
  scale_color_manual(
    values = c(`3` = "#29749E",
               `7` = "#122243")) +
    xlab("Score") +
    ylab("Sales Area")  +
    ggtitle("Score by Sales Area") +
  theme_minimal() +
  theme(plot.title = element_text(face = "bold", hjust = 0.5))
v2 
```


